Assessment of the KLGO Coastal Waterbird Survey Monitoring Program

### 2019-04-26 analysis/paper/paper.Rmd
ID'd next milestones for Madeleine to focus on (listed in intro bullets).

```
# NOTES To Madeleine
- see edits in survey_duration.R to aid interpretation by the viewer (e.g., changes in x axis tick locations, etc.)
- documentation in scripts

@@

# Issues to Resolve with content (narrative, figures, analyses)
- **Joel & Madeleine** will discuss survey_dates.R commands and how to modify them so that we get a sense of when surveys were & when target species was first (or last) seen each year. E.g., separate out any filtering process stemming from survey effort levels and timing. (thinking some thing like showing open circles for survey events PRIOR to first sighting...)
- Need to think about way to winnow out target species that do not occurr w/ adequate frequency to report on. E.g., those only seen <3 or 5 years??? Look at phenology graphs and determine threshold to apply. Consider some species N-ward and S-ward migration routes may differ.
- What is up with 2008 surveys? why such distinct timing?
- Think about Covariate EDA, if any...change in observers?
- Begin thinking about (Freq of Occurrence, Phenology, Abundance | target species) and then Richness - both Phenology (Day of Season) and across years.
- clean up structure of survey design section - consistent pattern in {Element Definition, Settings, Assessment}; as cleanly as possible separate space and time aspects of survey design (& effort) from spp-specific responses.

@@

#### Detection issues associated with survey conditions
.species specific counts w/in season by weather conditions and wave conditions
.species occurrence by wave conditions & weather conditions
.species age and composition by wave conditions & weather conditions 
.total # species by weather conditions, wave conditions; internally consistent contrasts to demonstrate value, or lack thereof, of current covariates.
.richness by major weather conditions (Northerlies, Southerlies, etc.) and for each spp
.distribution of weather conditions
.distribution of survey time by weather conditions
.diagnostic to ID key survey times to ensure high quality observers?
.consistency in ‘breeding periods’ across years (as captured in data)
.quick diagnostics to ID covariates seldom recorded, not consistently recorded, consistently valued (e.g., no basis for contrast), etc.

>>>>

```
# NOTES To Madeleine
- see edits in survey_duration.R to aid interpretation by the viewer (e.g., changes in x axis tick locations, etc.)
- documentation in scripts
- **NEXT Priority Milestones**: note - all references to line numbers are prior to any other edits!
- finish graphs of first & last observation per year for each target species (around line 386);**ultimately** add a fitted linear trend line to each of those (we need to account for frequency of surveying, so some subtle effort things here, but can resolve that later);
- graphs of species occurrence (around line 407) for each of target species; 
- covariate distribution graphs (lots of univariate & bivariate things; around line 534);
- plot of total # spp sighted each season across years (around line 600)

@@

# Issues to Resolve with content (narrative, figures, analyses)
- **Joel & Madeleine** will discuss survey_dates.R commands and how to modify them so that we get a sense of when surveys were & when target species was first (or last) seen each year. E.g., separate out any filtering process stemming from survey effort levels and timing. (thinking some thing like showing open circles for survey events PRIOR to first sighting...)
- Need to think about way to winnow out target species that do not occurr w/ adequate frequency to report on. E.g., those only seen <3 or 5 years??? Look at phenology graphs and determine threshold to apply. Consider some species N-ward and S-ward migration routes may differ.
- Joel will finish script uploading AOU data csv object & using to order Target Species
- What is up with 2008 surveys? why such distinct timing?
- Think about Covariate EDA, if any...change in observers?
- Begin thinking about (Freq of Occurrence, Phenology, Abundance | target species) and then Richness - both Phenology (Day of Season) and across years.
- clean up structure of survey design section - consistent pattern in {Element Definition, Settings, Assessment}; as cleanly as possible separate space and time aspects of survey design (& effort) from spp-specific responses.

@@

#### Detection issues associated with survey conditions
>MADELIENE: Next couple plots are priority.

.distribution of weather conditions
.distribution of survey time by weather conditions
.quick diagnostics to ID covariates seldom recorded, not consistently recorded, consistently valued (e.g., no basis for contrast), etc.
.species specific counts w/in season by weather conditions and wave conditions
.species occurrence by wave conditions & weather conditions
.species age and composition by wave conditions & weather conditions
.total # species by weather conditions, wave conditions; internally consistent contrasts to demonstrate value, or lack thereof, of current covariates.
.richness by major weather conditions (Northerlies, Southerlies, etc.) and for each spp
.diagnostic to ID key survey times to ensure high quality observers?
.consistency in ‘breeding periods’ across years (as captured in data)
